---
title: 'Statistical report: Interaction between HF treatment dosage and blood pressure for clinical outcome'
subtitle: 'DRAFT'
author: 'Statistician: Lina Benson'
  
date: "`r Sys.Date()`"
output:
  pdf_document:
    fig_caption: yes
    fig_height: 7
    fig_width: 7
    number_sections: yes
link-citations: yes
bibliography: references.bib
nocite: '@*'
header-includes:
   - \usepackage{draftwatermark}
---

\newpage 
\tableofcontents 
\listoffigures
\listoftables
\newpage


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, include = TRUE, comment = "",
  warning = FALSE, message = FALSE, fig.pos = "H",
  fig.path = "../output/figs/"
)
options(knitr.kable.NA = "")
```

```{r adjust_directory_if_needed, include=FALSE}
# Uncomment lines below if rmd file is placed in a subdirectory
knitr::opts_knit$set(root.dir = normalizePath("../"))
```

```{r load_project}
# 1. Set options in config/global.dcf
# 2. Load packages listed in config/global.dcf
# 3. Import functions and code in lib directory

ProjectTemplate::reload.project()

cacheon <- TRUE
```             

# Data handling

## Data source

SHFDB3, https://kiheartfailure.github.io/shfdb3/, v 3.1.5. 

## Inclusion/exclusion criteria

```{r flow}
mykable(flow, caption = "Flowchart")
```

First patient in: `r min(pdata$shf_indexdtm)` and last patient in: 
`r max(pdata$shf_indexdtm)`. 
Median age (IQR): 
`r pdata %>% summarise(med = dF(median(shf_age), dig = 1),
                       q1 = dF(quantile(shf_age, probs = 0.25), dig = 1),
                       q3 = dF(quantile(shf_age, probs = 0.75), dig = 1)) %>%
              mutate(out = paste0(med, " (", q1, "-", q3, ")")) %>%
              pull(out)` and 
`r pdata %>% count(shf_sex) %>%
  mutate(perc = dF(n / sum(n) * 100, 1)) %>%
  filter(shf_sex == "Female") %>%
  pull(perc)`% females.
  
## Created variables 

```{r outcodes}
mykable(metaout, caption = "Additional outcomes from NPR/Cause of Death Registry")
```

The supremum treatment variable is created as follows:

```{r codesuptret, code = readLines("../munge/02-suptreatvar.R"), echo = TRUE, eval = FALSE}

```

Note that mra is assumed to give 1 point if receiving mra at all since mra dose was 
not collected in the registry until 2014. 

Throughout the doses presented are % of target dose.

# Statistical analysis 

## General

All analyses were performed using `r sessionInfo()$R.version$version.string` [@r]. 
The level of significance is set to 5%, two-sided. No adjustment for multiple 
comparisons were made and therefore the results should be viewed with some care.

## Missing data

Missing data was imputed with multiple imputation using mice [@mice] for 
10 dataset and 10 iterations. Variables included in the model are indicated 
in Table \ref{tab:tab1}. The primary outcome, CVD/First HFH (cardiovascular 
death or first heart failure hospitalization), was included 
as the Nelson-Aalen estimator.

## Baseline characteristics

```{r, child = "../src/tab1.Rmd"}

```

```{r, child = "../src/tab_treat.Rmd"}

```

## Systolic blood pressure cut-offs

The functional form of systolic blood pressure modelled with a natural cubic spline 
with knots placed at the quintiles of sbp in a cox regression is plotted. 
Partly for the primary outcome, cvd/1 hf hosp, and partly for cvd as outcome. 
The curve is centered at the median sys bp. 

```{r, child = "../src/spline_ass_out_sbp.Rmd"}

```

\clearpage
\newpage

## Outcomes

Kaplan-Meier curves for CVD/First HFH are presented for the supremum treatment variable, ras, bbl and
mra. 

Crude and adjusted cox proportional hazards regressions were used to evaluate the association 
between the supremum dose variable/systolic blood pressure and the respective outcomes. 
Adjustment is performed for the variables indicated in Table \ref{tab:tab1}.

In order to present the supremum dose variable and systolic blood pressure 
in subgroups they are modelled with an interaction term. 
The HR (95% CI) presented is for continuous systolic blood pressure in units of 10
in the supremum treat subgroups and the continuous supremum treat variable in the 
systolic blood pressure subgroups (so the HR are from separate models, with the 
categorical and continuous SBP and treat interchanging).  

In the models with the supremum treatment variable ras, bbl and mra are not 
included as separate variables in the models. For the individual medications 
0 = no medication, 0.5 = <50% of target dose, 1 = >= 50% of target dose. 
For MRA 0 = no medication, 1 = medication since MRA dose was not 
collected in SwedeHF prior to 2014. A consistency analysis was therefore 
performed for the primary endpoint with only 
observations from 2014 with mra dose included in the supremum treatment variable. 

Data were censored at 2018-12-31 or death/emigration, 
so this assumes that the risk of hf hospitalization 
would have been the same for a censored patient had the censoring not occurred. 

### Assumptions

To be checked...

```{r, child = "../src/tab_event.Rmd"}

```

\clearpage
\newpage

```{r, child = "../src/km.Rmd"}

```

\clearpage
\newpage

```{r, child = "../src/tab_eventmod.Rmd"}

```

\clearpage
\newpage

Forest plot presenting the HR and 95% CI from a cox model with the variable  
systolic bp * sumpremum treatment. Both are categorical (in the table one is 
continuous). The group with high bp and high treatment is the reference. 

```{r, child = "../src/fig_hr.Rmd"}

```

\clearpage
\newpage

## Changes in SBP and HF medication dose (objective C)

The patients with a prior visit fulfilling the inclusion and none of the exclusion 
criteria were selected. 

A change variable was created defined as: 

- Stable group = Decrease in SBP between second last and last visit with less 
than 10 mmHg.

- Decrease SBP + increase treat = Decrease in SBP between second last and last 
visit with 10 mmHg or more and an increase in the supremum treatment variable.

- Decrease SBP + stable treat = Decrease in SBP between second last and last 
visit with 10 mmHg or more and a stable and decrease in the supremum treatment variable.

The baseline variables presented in the table and adjusted for in the models 
are from the LAST visit. 

```{r}
pdata_change <- pdata %>%
  filter(!is.na(treatsbp_diff))
```

```{r, child = "../src/histogram_change.Rmd"}

```

```{r, child = "../src/tab1_change.Rmd"}

```

```{r, child = "../src/tab_eventmod_change.Rmd"}

```

\clearpage
\newpage

# Reproducibility

## R session information {#sessioninfo}

```{r sessinfo}
sessionInfo()
```

## R code

The R code for all data handling and statistical analyses will be found 
https://github.com/KIHeartFailure/HFtreatSBP on publication. It is possible to
link to it from the resulting paper for increased transparency and code sharing. 
No data or output is stored in the repository. 

# References
